Typing a linear \picalc

### 2020-03-09 paper/paper.tex
Add bits on structural congruence and reduction

\subsection{Structural congruence}

Structural congruence is a congruent equivalence relation on processes.
Any two structurally congruent processes are strongly bisimilar: they are can follow each other's reduction steps \cite{}.
Figure \ref{struct-cong-base} lists the base cases of structural congruence.

@@

  \inferrule
  {eq_1 P =_r Q \\ \; eq_2 Q =_s R}
  {\constr{trans} \; eq_1 \; eq_2 : P =_{\constr{two} \; r \; s} R}
\end{mathpar}
\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.
  Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred}
\label{struct-cong}
\end{figure}

\subsection{Reduction}

@@

  \inferrule
  { }
  {\constr{nothing} : \Channel_n}

  \inferrule
  {i : \Var_n}
  {\constr{just} \; i : \Channel_n}

@@

  \inferrule
  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}

@@

  \inferrule
  {eq : P = P' \\ red : P' \reduce_i Q}
  {\constr{struct} \; eq \; red : P \reduce_i Q}
\end{mathpar}
\caption{Operational semantics indexed by reducing channel}
\end{figure}

keeping track of the variable on which communication occurs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Linear typing rules}

@@

\subsubsection{Example type systems}

@@

\subsection{Variable references}

@@

\subsection{Typing with leftovers}

\subsubsection{Typing relation}

@@

  \inferrule
  {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
   r : \types{\gamma}{\Delta}{Q}{\Xi}}
  {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
\end{mathpar}
\caption{This is...}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject reduction}\label{subject-reduction}

@@

\subsection{Subject congruence}\label{subject-congruence}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Related work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future work}

>>>>

\subsection{Structural Congruence}

Structural congruence is a congruent equivalence relation on processes.
Any two structurally congruent processes are strongly bisimilar: they are can follow each other's reduction steps \cite{}.
Figure \ref{struct-cong-base} lists the base cases of structural congruence.
The type $\Unused_0 \; Q$ witnesses that index $0$ does not appear nor in the inputs nor in the outputs of process $Q$ --- it does so by traversing $Q$.
The function $lower \; 0 \; Q \; uQ$ traverses $Q$ decrementing every index bigger than $0$.
Finally, $swap \; 0 \; P$ traverses $P$ (scoped under $\suc \suc n$) and swaps variable references $0$ and $\suc 0$.

@@

  \inferrule
  {eq_1 : P =_r Q \\ \; eq_2 : Q =_s R}
  {\constr{trans} \; eq_1 \; eq_2 : P =_{\constr{two} \; r \; s} R}
\end{mathpar}
\caption{Structural rewriting rules lifted to a congruent equivalence relation indexed by a recursion tree.
  Premises $P$, $P'$, $Q$, and $R$ are of type $\Process_n$ where $n$ can be inferred.}
\label{struct-cong}
\end{figure}

\subsection{Operational Semantics}

Figure \ref{reduction} models the operational semantics of the \picalc{}.
Processes put in parallel reduce when they communicate over a common variable.
The receiving process substitutes references to its most immediate variable with references to the variable sent by the process doing the output, and is then \emph{lowered} --- all variable references are decreased by one.
Reduction is closed under structural congruence and goes under parallel composition and scope restriction
--- but notably not under input nor output, otherwise the sequencing of actions would not be preserved.
In \S \ref{subject-reduction} we prove that if $P$ reduces to $Q$ and $P$ is well-typed, so is $Q$.
However, this reduction operation is effectful: it consumes the variable over which communication happens.
If this variable is external to $P$ (it resides in its context) then the context in which $Q$ is typed must change.
To keep track of this information, we lift the variable index over which communication occurs to the type level.
Every time we come out of a binder we decrement this variable.
To do so, we make use of a $dec$ function that saturates at $\constr{internal}$ --- as opposed to an $inc$ function that would not.

@@

  \inferrule
  { }
  {\constr{internal} : \Channel_n}

  \inferrule
  {i : \Var_n}
  {\constr{external} \; i : \Channel_n}

@@

  \inferrule
  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{external \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}

@@

  \inferrule
  {eq : P = P' \\ red : P' \reduce_i Q}
  {\constr{struct} \; eq \; red : P \reduce_i Q}
\end{mathpar}
\caption{Operational semantics indexed by the channel over which reduction occurs.}
\label{reduction}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Resource-aware Typing System}

@@

\subsubsection{Example Type Systems}

@@

@@

\subsection{Typing with Leftovers}

@@

  \inferrule
  {l : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
   r : \types{\gamma}{\Delta}{Q}{\Xi}}
  {\constr{comp} \; l \; r : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
\end{mathpar}
\caption{This is...}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject Reduction}\label{subject-reduction}

@@

\subsection{Subject Congruence}\label{subject-congruence}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Related Work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}
